BARBELL LIFTS MODEL
========================================================

***Synopsys***

This study is conducted to determine how the barbell lifts are done, based on data get from accelerometers on the belt, forearm, arm, and dumbell.

This "classification model" was tested using decission trees and random forest (too much time processing data on a "MAC 8gb ram and core i7"). 

The training data was split in Training and Testing data. Training 70%, Testing 30%. 

The cross validation was conducted using three random samples, and the out of sample error is estimated by average of three models.

The model reports 66.17% accuracy average on testing data (extracted from training dataset):
Model 1: 66.15%
Model 2: 66.17%
Model 3: 66.19%

Out of sample error expected =  33%

**Loading data and Exploratory Analysis**


```{r}
library (rattle)
library(AppliedPredictiveModeling)
library(caret)
library(Hmisc)
library(e1071)
library(psych)

setwd("~/Coursera/Data Science/Practical Machine Learning/project1")
trainingSetOR <- read.csv("pml-training.csv")
testingSetOR <- read.csv("pml-testing.csv")

trainingSet <- trainingSetOR[,-c(12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,50,51,52,53,54,55,56,57,58,59,69,70,71,72,73,74,75,76,77,78,79,80,81,82,83,87,88,89,90,91,92,93,94,95,96,97,98,99,100,101,103,104,105,106,107,108,109,110,111,112,125,126,127,128,129,130,131,132,133,134,135,136,137,138,139,141,142,143,144,145,146,147,148,149,150)]

testingSet <- testingSetOR[,-c(12,13,14,15,16,17,18,19,20,21,22,23,24,25,26,27,28,29,30,31,32,33,34,35,36,50,51,52,53,54,55,56,57,58,59,69,70,71,72,73,74,75,76,77,78,79,80,81,82,83,87,88,89,90,91,92,93,94,95,96,97,98,99,100,101,103,104,105,106,107,108,109,110,111,112,125,126,127,128,129,130,131,132,133,134,135,136,137,138,139,141,142,143,144,145,146,147,148,149,150)]



#str(trainingSet)
#describe(trainingSet)
trainIndex = createDataPartition(y=trainingSet$classe, p = 0.70,list=FALSE) #70% for training
training = trainingSet[trainIndex,]
testing = trainingSet[-trainIndex,]

#RF<-train(classe ~ ., data=training, method="rf", prox=TRUE)

# Assign values to Clasee ... for ploting purposses
# A=1, B=2, C=3, D=4, E=5
trainingPlot <- training
trainingPlot$Num <- 0
colNum=61
trainingPlot[trainingPlot$classe=="A",colNum] <- 1
trainingPlot[trainingPlot$classe=="B",colNum] <- 2
trainingPlot[trainingPlot$classe=="C",colNum] <- 3
trainingPlot[trainingPlot$classe=="D",colNum] <- 4
trainingPlot[trainingPlot$classe=="E",colNum] <- 5
hist(trainingPlot$Num, main="Classe Histogram. A=1, B=2, C=3, D=4, E=5")
```
Little skewness at right. But assume normal distribution.

```{r}
qplot(trainingPlot$X,trainingPlot$Num, colour=trainingPlot$user_name, data=trainingPlot, main="Classe in Numbers: A=1, B=2, C=3, D=4, E=5")
```
X is an important variable as looks like each User were asked to do de exercise in a different way at by sequence in time.


Modelling:  3 random models for cross validation
===================
```{r}
# Model 1
modFit1 <- train(classe ~ ., method="rpart", data=training) #Commented for publishing results
confusionMatrix(testing$classe, predict(modFit1, newdata=testing))
```

** Model 1 Accuracy: 66.17% . Out of sample Error = 34% **

```{r}
# Model 2
trainIndex2 = createDataPartition(y=trainingSet$classe, p = 0.70,list=FALSE) #70% for training
training2 = trainingSet[trainIndex2,]
testing2 = trainingSet[-trainIndex2,]

modFit2 <- train(classe ~ ., method="rpart", data=training2)  #Commented for publishing results
confusionMatrix(testing2$classe, predict(modFit2, newdata=testing2))
```

**Model 2 Accuracy: 66.15% . Out of sample Error = 34% **

```{r}
# Model 3
trainIndex3 = createDataPartition(y=trainingSet$classe, p = 0.70,list=FALSE) #70% for training
training3 = trainingSet[trainIndex3,]
testing3 = trainingSet[-trainIndex3,]

modFit3 <- train(classe ~ ., method="rpart", data=training3)   #Commented for publishing results
confusionMatrix(testing3$classe, predict(modFit3, newdata=testing3))
```

**Model 3 Accuracy: 66.19%. Out of sample Error = 34%**


Applying model to test dataset
============
```{r}

results <- predict(modFit1, newdata=testingSet)

class(results)

answers <- as.vector(results, mode = "character")
is.vector(answers)
class(answers)


pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}


pml_write_files(answers)

```


